%% ----------------------------------------------------------------
%% Thesis.tex -- MAIN FILE (the one that you compile with LaTeX)
%% ---------------------------------------------------------------- 

% Set up the document
\documentclass[a4paper, 11pt, oneside]{Thesis}  % Use the "Thesis" style, based on the ECS Thesis style by Steve Gunn
\graphicspath{Figures/}  % Location of the graphics files (set up for graphics to be in PDF format)

% Include any extra LaTeX packages required
\usepackage[square, numbers, comma, sort&compress]{natbib}  % Use the "Natbib" style for the references in the Bibliography
\usepackage{verbatim}  % Needed for the "comment" environment to make LaTeX comments
\usepackage{vector}  % Allows "\bvec{}" and "\buvec{}" for "blackboard" style bold vectors in maths
\hypersetup{urlcolor=blue, colorlinks=true}  % Colours hyperlinks in blue, but this can be distracting if there are many links.
\makeglossaries
%% ----------------------------------------------------------------
\begin{document}
\frontmatter      % Begin Roman style (i, ii, iii, iv...) page numbering

% Set up the Title Page
\title  {Reverse Engineering Object-Oriented Systems into Umple: An Incremental and Rule-Based Approach}
\authors  {\texorpdfstring
            {\href{mgarzon@uottawa.ca}{Miguel Garz\'{o}n}}
            {Miguel A. Garz\'{o}n}
            }
\date  {\normalsize \copyright Miguel Garz\'{o}n, Ottawa, Canada, 2015}

\maketitle
%% ----------------------------------------------------------------

\setstretch{1.3}  % It is better to have smaller font and larger line spacing than the other way round

% Define the page headers using the FancyHdr package and set up for one-sided printing
\fancyhead{}  % Clears all page headers and footers

\rhead{\thepage}  % Sets the right side header to show the page number
\lhead{}  % Clears the left side page header

\pagestyle{fancy}  % Finally, use the "fancy" page style to implement the FancyHdr headers

%% ----------------------------------------------------------------
% The "Funny Quote Page"
%\pagestyle{empty}  % No headers or footers for the following pages
\setcounter{page}{2}
\null\vfill
% Now comes the "Funny Quote", written in italics
\textit{``Live as if you were to die tomorrow. Learn as if you were to live forever.''}

\begin{flushright}
Mahatma Gandhi   
\end{flushright}

\vfill\vfill\vfill\vfill\vfill\vfill\null
\clearpage  % Funny Quote page ended, start a new page
%% ----------------------------------------------------------------

% The Abstract Page
\addtotoc{Abstract}  % Add the "Abstract" page entry to the Contents
\abstract{
\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics

This thesis investigates a novel approach to reverse engineering, in which modeling information such as UML associations, state machines and attributes is incrementally added to code written in Java or C++, while maintaining the system in a textual format. Umple is a textual representation that blends modeling in UML with programming language code. The approach, called umplification, produces a program with behavior identical to the original one, but written in Umple and enhanced with model-level abstractions. As the resulting program is Umple code, our approach eliminates the distinction between code and model. We implemented automated umplification in  a tool called the Umplificator. The tool is rule-driven: code, including Umple code, is parsed and processed into an internal representation, which is then operated on by rules; transformed textual code and model, in Umple, is then generated. The rules used to transform code to model have been iteratively refined by using the tool on a variety of open-source software systems.

The thesis consists of three main parts. The first part (Chapters 1 and 2) present the research questions and research methodology, as well as introducing Umple and the background necessary to understand the rest of the thesis. The umplification method is presented at increasing levels of detail through Chapters 3 and 4. Chapters 5 and 6 present the tool and evaluation of our approach, respectively. An analysis of related work, and comparison to our own, appears in Chapter 7. Finally, conclusions and future work directions are presented in Chapter 8.
}

\clearpage  % Abstract ended, start a new page
%% ----------------------------------------------------------------

\setstretch{1.3}  % Reset the line-spacing to 1.3 for body text (if it has changed)

% The Acknowledgements page, for thanking everyone
\acknowledgements{
\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics

Foremost, I would like to express my sincere gratitude to my dear supervisor, Professor Timothy Lethbridge, for guiding me during my graduate studies. I greatly appreciate his vast knowledge, dedication and assistance in writing proposals, scholarship applications, papers and this thesis. Throughout my studies, Tim contributed to my development as a researcher, professor and as a person.

Besides my supervisor, I would also like to thank the members of my PhD committee for the feedback provided at the different stages or this research. 

My sincere thanks to Professors Daniel Amyot and Liam Peyton for the continuous encouragement to my work. More than my instructors, they have been mentors who helped shape my career path. 

I thank each of the members of the Complexity Reduction in Software Engineering (CRuiSE) research group. Particular and well deserved thanks to Hamoud Aljamaan for all of his constructive feedback.

I wish to deeply thank the University of Ottawa for being my second home for the past 10 years and for offering me all the support required to carry out my research work.

I would also like to thank my whole family for the support they provided me through my entire life. In particular, I would like to thank my mother Patricia, my brothers Laura and Sebastian for their love and continuous encouragement during this busy period of my life.

Last, but not least, I must express my gratitude to Idalia for her understanding and love during the past few years and for her contribution on the graphics of this thesis.
}
\clearpage  % End of the Acknowledgements
%% ----------------------------------------------------------------

\pagestyle{fancy}  %The page style headers have been "empty" all this time, now use the "fancy" headers as defined before to bring them back


%% ----------------------------------------------------------------
\lhead{\emph{Contents}}  % Set the left side page header to "Contents"
\tableofcontents  % Write out the Table of Contents

%% ----------------------------------------------------------------
\lhead{\emph{List of Figures}}  % Set the left side page header to "List if Figures"
\listoffigures  % Write out the List of Figures

%% ----------------------------------------------------------------
\lhead{\emph{List of Tables}}  % Set the left side page header to "List of Tables"
\listoftables  % Write out the List of Tables

%% ----------------------------------------------------------------
%\lhead{\emph{List of Listings}}  
%\lstlistoflistings  % Write out the List of Tables
%% ----------------------------------------------------------------
\setstretch{1.5}  % Set the line spacing to 1.5, this makes the following tables easier to read
\clearpage  % Start a new page
\lhead{\emph{Glossary}}  % Set the left side page header to "Abbreviations"
%% ----------------------------------------------------------------
\mainmatter	  % Begin normal, numeric (1,2,3...) page numbering
\pagestyle{fancy}  % Return the page headers back to the "fancy" style

% Include the chapters of the thesis, as separate files
% Just uncomment the lines as you write the chapters
\lhead{\emph{\rightmark}} 

\input{Chapters/Chapter1_Intro} % Introduction

\input{Chapters/Chapter2_Background} % Background  

\input{Chapters/Chapter3_Core} % Reverse Engineering

\input{Chapters/Chapter4_Detection} % Detection Mechanisms

\input{Chapters/Chapter5_Tool} % The Umplificator Technologies

\input{Chapters/Chapter6_Eval} % Evaluation

\input{Chapters/Chapter7_Related} % Related Work

\input{Chapters/Chapter8_Conclusions} % Conclusions and Contributions

%% ----------------------------------------------------------------
% Now begin the Appendices, including them as separate files

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
 
%\appendix % Cue to tell LaTeX that the following 'chapters' are Appendices

%\input{Appendices/AppendixA}	% Appendix Title

%\input{Appendices/AppendixB} % Appendix Title

%\input{Appendices/AppendixC} % Appendix Title

\addtocontents{toc}{\vspace{2em}}  % Add a gap in the Contents, for aesthetics
\backmatter

%% ----------------------------------------------------------------
\label{Bibliography}
\lhead{\emph{Bibliography}}  % Change the left side page header to "Bibliography"
\bibliographystyle{unsrtnat}  % Use the "unsrtnat" BibTeX style for formatting the Bibliography
\bibliography{Bibliography}  % The references (bibliography) information are stored in the file named "Bibliography.bib"

\end{document}  % The End
%% ----------------------------------------------------------------